Division and Factorization of p-Adic Polynomials

### 2016-01-12 polynomials/muldivfact.tex
Drawing in main proof
Continued.
Tristan's figure is now Figure 2
Abstract added

\begin{abstract}
\end{abstract}

@@

Noting that $A_0$ and $P$ agree up to degree $d$ and that $\NP(P)$ has 
an extremal point at abscissa $d$, we get $\NF(A_0) = \varphi_{|[0,d]}$.
Clearly $b_1(V_0) \geq 0$ since $V_0 = 1$. It follows from the 
definitions that $R_0 = S_0 = P \mod A_0 = (P{-}A_0) \mod A_0$. 
We remark that $P{-}A_0 = \sum_{i=d+1}^n a_i X^i.$
Using Theorem \ref{theo:EDivisionNP}, we obtain that
$b_\varphi(P{-}A_0) \geq \kappa$ and then $b_\varphi(R_0) = b_\varphi(S_0)
\geq \kappa$. 
%Mettre le dessin ici
%
%
%

\begin{tikzpicture}
[xscale=0.7, yscale=0.5]
\fill[blue!30] (4,9.5)--(4,1.6)--(2,1.2)--(0.5,2.8)--(0,6.8)--(0,9.5)--cycle;
\fill[yellow!50] (6,9.5)--(6,2)--(4,1.6)--(4,9.5)--cycle;
\fill[green!30] (6,9.5)--(6,2)--(8,4)--(11,9)--(11,9.5)--cycle;
\draw[->] (-0.5,0)--(11.5,0);
\draw[->] (0,-0.5)--(0,9.5);
\draw[very thick] (0,6)--(0.5,2)--(2,.4)--(5,1);
\draw(5,1)--(8,4)--(11,9);
\draw[thick] (4,9.5)-- (4,1.6)--(2,1.2)--(0.5,2.8)--(0,6.8);
\draw[dotted] (6,2)--(4,1.6);
%useless part?
\draw[dotted] (7,3)--(2,2)--(0.5,3.6)--(0,7.6);
\draw[dotted] (8,4)--(2,2.8)--(0.5,4.4)--(0,8.4);
\draw[dotted] (9,5.667)--(2,4.267)--(0.5,5.867)--(0.046,9.5); %%(0,9.867);
\draw[dotted] (10,7.333)--(2,5.733)--(0.5,7.333)--(0.229,9.5); %%--(0,11.333);
\draw[dotted] (11,9)--(2,7.2)--(0.5,8.8)-- (.413,9.5);%%(0,12.8);

\draw[<->] (4,.8)--(4,1.6);
\draw[dotted,thick] (5,0)--(5,9.5);
\draw[dotted,thick] (4,0)--(4,9.5);
\draw[dotted,thick] (6,0)--(6,9.5);
\node[scale=0.8, below left] at (0,0) { $0$ };
\node[scale=0.8, below] at (5,0) { $d$ };
\node[scale=0.8, below] at (4,0) { $d{-}1$ };
\node[scale=0.8, below] at (6,0) { $d{+}1$ };
\node[scale=0.9, below right] at (2.5,.6) { $\mathbf{A_0}$ };
\node[scale=0.9, right] at (10.2,7.3) { $P$ };
%\node[scale=0.9, right] at (10.5,4.4) { $\Delta$ };
\node[scale=0.9] at (4.2,1.2) { $\kappa$ };
\node[thick] at (3,9) { $R_0$ };
\node[thick] at (3,8.5) { $S_0$ };
%\draw[dashed] (11,5)--(8,4);
%\draw (8,4)--(2,2)--(0.5,4)--(0,8);
\end{tikzpicture}

Finally observe that:
$$T_0 = (1 - B_0) \mod A_0 = \big((A_0 - P) \div A_0\big) \mod A_0.$$
Therefore $b_0(T_0) \geq \kappa$ results from $b_\varphi(A{-}P_0) \geq 
\kappa$ thanks to Lemma~\ref{lem:bphib0}. 
We have then established~\eqref{eq:induction} when $i = 0$.

>>>>

\begin{abstract}
We study two important operations on polynomials defined over complete 
discrete valuation fields (\emph{e.g.} $\Q_p$ or $k((t))$ for a 
field $k$), namely Euclidean division and factorization. We consider 
notably the question of computing slope factorizations and design a new 
simple and efficient algorithm based on Newton iteration to solve it.
One of its main feature consists in avoiding working over an extension 
of the base field.
We constantly pay particular attention to stability: we design several
precision models, compare them and always discuss optimality.
\end{abstract}

@@

Noting that $A_0$ and $P$ agree up to degree $d$ and that $\NP(P)$ has 
an extremal point at abscissa $d$, we get $\NF(A_0) = \varphi_{|[0,d]}$.
Clearly $b_1(V_0) \geq 0$ since $V_0 = 1$. It follows from the 
definitions that $R_0 = S_0 = P \mod A_0 = (P{-}A_0) \mod A_0$. 
We remark that $P{-}A_0 = \sum_{i=d+1}^n a_i X^i.$
Using Theorem \ref{theo:EDivisionNP}, we obtain that
$b_\varphi(P{-}A_0) \geq \kappa$ and then $b_\varphi(R_0) = b_\varphi(S_0)
\geq \kappa$ (see Figure~\ref{fig:boundR0S0}).
%
\begin{figure}
\hfill
\begin{tikzpicture}[xscale=0.7, yscale=0.5]
\fill[blue!30] (4,9.5)--(4,1.6)--(2,1.2)--(0.5,2.8)--(0,6.8)--(0,9.5)--cycle;
\fill[yellow!50] (6,9.5)--(6,2)--(4,1.6)--(4,9.5)--cycle;
\fill[green!30] (6,9.5)--(6,2)--(8,4)--(11,9)--(11,9.5)--cycle;
\draw[->] (-0.5,0)--(11,0);
\draw[->] (0,-0.5)--(0,9.5);
\draw[very thick] (0,6)--(0.5,2)--(2,.4)--(5,1);
\draw(5,1)--(8,4)--(11,9);
\draw[thick] (4,9.5)-- (4,1.6)--(2,1.2)--(0.5,2.8)--(0,6.8);
\draw[dotted] (6,2)--(4,1.6);
%useless part?
\draw[dotted] (7,3)--(2,2)--(0.5,3.6)--(0,7.6);
\draw[dotted] (8,4)--(2,2.8)--(0.5,4.4)--(0,8.4);
\draw[dotted] (9,5.667)--(2,4.267)--(0.5,5.867)--(0.046,9.5); %%(0,9.867);
\draw[dotted] (10,7.333)--(2,5.733)--(0.5,7.333)--(0.229,9.5); %%--(0,11.333);
\draw[dotted] (11,9)--(2,7.2)--(0.5,8.8)-- (.413,9.5);%%(0,12.8);
\draw[very thick,dotted] (5,1)--(11,2.2);
\draw[dotted] (0,0)-- (5,1);

\draw[<->] (4,.8)--(4,1.6);
\draw[dotted,thick] (5,0)--(5,9.5);
\draw[dotted,thick] (4,0)--(4,9.5);
\draw[dotted,thick] (6,0)--(6,9.5);
\node[scale=0.8, below left] at (0,0) { $0$ };
\node[scale=0.8, below] at (5,0) { $d$ };
\node[scale=0.8, below] at (4,0) { $d{-}1$ };
\node[scale=0.8, below] at (6,0) { $d{+}1$ };
\node[scale=0.85, below left] at (1.5,1.3) { $\mathbf{A_0}$ };
\node[scale=0.85, right] at (10.1,7.3) { $NF(P)$ };
%\node[scale=0.9, right] at (10.5,4.4) { $\Delta$ };
\node[scale=0.9] at (4.2,1.2) { $\kappa$ };
\node[scale=0.8] at (3,9) { $NP(R_0)$ };
\node[scale=0.8] at (3,8.3) { $NP(S_0)$ };
\node[right,scale=0.9,rotate=35] at (6.2,1.7) { $\text{slope} = \lambda_1$ };
\node[right,scale=0.9,rotate=8] at (6.8,.9) { $\text{slope} = \lambda_0$ };
\node[thick] at (10,1.6) { $\varphi$ };
%\draw[dashed] (11,5)--(8,4);
%\draw (8,4)--(2,2)--(0.5,4)--(0,8);
\end{tikzpicture}
\hfill\null

\caption{Bound on $\NF(R_0)$ and $\NF(S_0)$}
\label{fig:boundR0S0}
\end{figure}
%
Finally observe that:
$$T_0 = (1 - B_0) \mod A_0 = \big((A_0 - P) \div A_0\big) \mod A_0.$$
Therefore $b_0(T_0) \geq \kappa$ results from $b_\varphi(A{-}P_0) \geq 
\kappa$ thanks to Lemma~\ref{lem:bphib0}. 
We have then established~\eqref{eq:induction} when $i = 0$.
